---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "50%"
)
library(adas.utils)
library(magrittr)
```

# adas.utils

<!-- badges: start -->
<!-- badges: end -->

The goal of adas.utils is to provide some utility functions to be used in the course *Analysis of Data and Statistics* at the University of Trento, Italy. Course contents are on <https://paolobosetti.quarto.pub/ADAS/>.


## Installation

You can install the development version of adas.utils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pbosetti/adas.utils", build_vignettes = TRUE)
```

## Examples

### Loading data sets

On <https://paolobosetti.quarto.pub/data.html> there are a list of example datasets to be used during the course. You can load them with the `examples_url` function:

```{r}
examples_url("battery.dat") |> read.table(header=T) |> head()
```

### Chauvenet's criterion

The Chauvenet's criterion is a method to identify possible outliers in a sample. Here is an example:

```{r}
x <- rnorm(100)
x[50] <- 10
chauvenet(x)
```

### Daniel's plot

Daniel's plot is a QQ plot of the effects of a non-replicated factorial model. Here is an example:

```{r}
daniel_plot_qq(lm(Y~A*B*C*D, data=filtration))
```
### Tukey's test plot

The `stats::TukeyHSD` function in base R provides a way to perform multiple comparisons of means. It can print the results in tabular form or can be plotted with the `plot` method:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
examples_url("battery.dat") %>%
  read_table() %>%
  mutate(across(c(Temperature, Material), factor)) %>%
  mutate(Material = LETTERS[Material]) %>% 
  ggTukey(Response~Temperature, splt=~Material)
```

### Pareto chart

The `pareto_chart` function is a generic function that creates a Pareto chart either from a general data frame or from the effects of a `lm` object. Here is an example:

```{r}
library(tidyverse)
set.seed(1)
tibble(
  val=rnorm(10, sd=5),
  cat=LETTERS[1:length(val)]
  ) %>%
  pareto_chart(labels=cat, values=val)

# For a linear model:
pareto_chart(lm(Y~A*B*C*D, data=filtration))
```

### Design of Experiments

There is a host of functions to be used in the context of Design of Experiments. Here is an example to prepare a design matrix for a $2^{5-2}$ fractional factorial plan:

```{r}
fp_design_matrix(5) %>% 
  fp_fraction(~A*B*C*D) %>% 
  fp_fraction(~B*C*D*E)
```

You can also prepare a design matrix for a $2^n$ full factorial plan, and later on augment it with a central point and then to a central composite design with axial points:

```{r}
fp_design_matrix(3, rep=2) %>% 
  fp_augment_center(rep=5) %>% 
  fp_augment_axial(rep=2)
```




# Author

Paolo Bosetti, University of Trento
