---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(adas.utils)
```

# adas.utils

<!-- badges: start -->
<!-- badges: end -->

The goal of adas.utils is to provide some utility functions to be used in the course *Analysis of Data and Statistics* at the University of Trento, Italy. Course contents are on <https://paolobosetti.quarto.pub/ADAS/>.


## Installation

You can install the development version of adas.utils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pbosetti/adas.utils")
```

## Examples

### Loading data sets

On <https://paolobosetti.quarto.pub/data.html> there are a list of example datasets to be used during the course. You can load them with the `examples_url` function:

```{r}
examples_url("battery.dat") |> read.table(header=T) |> head()
```

### Chauvenet's criterion

The Chauvenet's criterion is a method to identify possible outliers in a sample. Here is an example:

```{r}
x <- rnorm(100)
x[50] <- 10
chauvenet(x)
```

### Daniel's plot

Daniel's plot is a QQ plot of the effects of a non-replicated factorial model. Here is an example:

```{r}
daniel_plot(lm(Y~A*B*C*D, data=filtration))
```

### Pareto chart

The `pareto_chart` function is a generic function taht creates a Pareto chart either from a general data frame or from the effects of a `lm` object. Here is an example:

```{r}
library(tidyverse)
set.seed(1)
tibble(
  val=rnorm(10, sd=5),
  cat=LETTERS[1:length(val)]
  ) %>%
  pareto_chart(labels=cat, values=val)

# For a linear model:
pareto_chart(lm(Y~A*B*C*D, data=filtration))
```

