---
title: "ADAS Utils vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ADAS Utils vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(adas.utils)
library(tidyverse)
```

# Factorial plans

The package provides tools for dealing with factorial plan according to the *Design of Experiments* (DoE) protocols. The functions for dealing with DoE have names starting with `fp_`. As much as possible, we are aiming at a tidyverse-like syntax, so that the functions can be used in a pipe.

## Full factorial plan

You can create a full factorial plan with the `fp_design_matrix` function, passing the number of factors:
```{r}
(dm <- fp_design_matrix(2, rep=2) %>% 
  mutate(Y=rnorm(n())))
```

In this case, the factors are the first `n` capital letters.

If you want different factor names, use a right-side-only formula combining all the named factors with `*`:

```{r}
fp_design_matrix(~Speed*Weight)
```

The yield column `Y` must then be completed according to the randomized `RunOrder` column.

## Custom levels

If you want a $k^n$ factorial plan with custom levels, pass the `levels` argument. In this case, though, the `.treat` column with Yate's treatment codes would be `NA`:

```{r}
fp_design_matrix(2, levels=-1:1)
```

## Augment a plan

You can augment a plan by adding a central point, typically repeated:

```{r}
fp_design_matrix(3) %>%
  fp_augment_center(rep=4)
```

Then if needed (because the analysis show low p-value for the quadratic term) you can add axial points to get a central composite design:

```{r}
fp_design_matrix(3) %>% 
  fp_augment_center(rep=3) %>% 
  fp_augment_axial(rep=2)
```



## Fractional factorial plan

A full $2^n$ factorial plan can be then reduced to a fractional factorial plan $2^{n-p}$ by applying the `fp_fraction` function $p$ times. For example, to get a $2^{5-2}$ plan with the defining relationships $I=ABCD$ and $I=BCDE$:

```{r}
fp_design_matrix(5) %>% 
  fp_fraction(~A*B*C*D) %>% 
  fp_fraction(~B*C*D*E)
```

Note that with the `remove` option you can control if you want to keep both fractions, and later on `filter(ABC==1)` them out.

```{r}
fp_design_matrix(3) %>% 
  fp_fraction(~A*B*C, remove=FALSE)
```


## Alias structure

The functions `fp_alias` and `fp_alias_list` can be used to generate the alias structure of a fractional factorial plan. **This functionalities are currently under development**


# Statistics

The package provides some useful functions for statistical analysys of data.

## Plotting

### Normal probability plot

The normal probability plot is provided as an alternative to the quantile-quantile plot:

```{r warning=FALSE}
df <- tibble(
  xn = rnorm(100, mean=20, sd=5),
  xu = runif(100, min=0, max=40)
)

df %>% normplot(xn)
df %>% normplot(xu)
```

### Pareto chart

The Pareto chart is a bar chart that displays the relative importance of problems in a format that is very easy to interpret. The bars are sorted in descending order, and the cumulative percentage of the total is shown by the line.

It can prove useful in the context of factorial plans, to identify the most important factors, or in sensityvity analysis, to identify the most important parameters.

The package provides a *generic function*, `pareto_chart`, that can be used with a tibble (or a data frame), or with a linear model (an `lm` object). In the latter case, the function produces the Pareto chart of the model effects.

For the general case, when you have a tibble with values and names:

```{r}
set.seed(1)
tibble(
  val=rnorm(10, sd=5),
  cat=LETTERS[1:length(val)]
  ) %>%
  pareto_chart(labels=cat, values=val)
```

For the case of a linear model:

```{r}
filtration %>% 
  lm(Y~A*B*C*D, data=.) %>%
  pareto_chart()
```

### Daniel's plot

In case of non-replicated factorial plans, the Daniel's plot can be used to identify the most important factors: a quantile-quantile plot of the factors effects shows the significant factors and interactions off the diagonal.

```{r}
daniel_plot_qq(lm(Y~A*B*C*D, data=filtration))
```

If you prefer, you can rather use a half-normal plot:

```{r}
filtration %>% 
  lm(Y~A*B*C*D, data=.) %>%
  daniel_plot_hn(nlab=6, repel=TRUE)

```


It shows that none of the effects containing the `B` factor are significant, so we can reduce the model to `Y~A*C*D`:

```{r}
filtration %>% 
  lm(Y~A*C*D, data=.) %>%
  anova()
```

Even better, the model can be further reduced to `Y~A*C+A*D`. Compare this conclusion with the last Pareto chart above.


# Utilities

This package also provides a function for easily loading data files made available on the accompanying course documentation on <https://paolobosetti.quarto.pub/data>:

```{r}
examples_url("battery.dat") %>%  read.table(header=TRUE)
```

