---
title: "ADAS Utils vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(adas.utils)
library(tidyverse)
```

# Factorial plans

The package provides tools for dealing with factorial plan according to the Design of Experiments protocols.

## Full factorial plan

You can create a full factorial plan with the `fp_design_matrix` function, passing the number of factors:
```{r}
(dm <- fp_design_matrix(2, rep=2) %>% 
  mutate(Y=rnorm(n())))
```

In this case, the factors are the first `n` capital letters.

If you want different factor names, use a formula combining all the named factors with `*`:

```{r}
fp_design_matrix(~Speed*Weight)
```

The yield column `Y` must then be completed according to the randomized `RunOrder` column.

## Custom levels

If you want a $k^n$ factorial plan with custom levels, pass the `levels` argument. In this case, though, the `.treat` column with Yate's treatment codes would be `NA`:

```{r}
fp_design_matrix(2, levels=-1:1)
```

## Augment a plan

you can augment a plan by adding a central point, typically repeated:

```{r}
fp_design_matrix(2, rep=2) %>% 
  fp_augment_center()
```


## Fractional factorial plan

A full $2^n$ factorial plan can be then reduced to a fractional factorial plan $2^{n-p}$ by applying the `fp_fraction` function $p$ times. For example, to get a $2^{5-2}$ plan with the defining relationships $I=ABCD$ and $I=BCDE$:

```{r}
fp_design_matrix(5) %>% 
  fp_fraction(~A*B*C*D) %>% 
  fp_fraction(~B*C*D*E)
```

Note that with the `remove` option you can control if you want to keep both fractions, and later on `filter(ABC==1)` them out.

```{r}
fp_design_matrix(3) %>% 
  fp_fraction(~A*B*C, remove=FALSE)
```


## Alias structure

The functions `fp_alias` and `fp_alias_list` can be used to generate the alias structure of a fractional factorial plan. **This functionalities are currently under development**
